Modeling the Circumnuclear Mediums and Accetion Rates of Quiescent Supermassive Black Holes

### 2014-11-11 ms.tex
Changing appendix around.

\newcommand{\dxdy}[2]{\frac{\partial #1}{\partial #2} }
\newcommand{\ddr}[1]{\dxdy{#1}{r}}
\newcommand{\drhodt}{\dxdy{\rho}{t}}
\newcommand{\dpdr}{\dxdy{p}{r}}
\newcommand{\dvdr}{\dxdy{v}{r}}
\newcommand{\dsdr}{\dxdy{s}{r}}
\newcommand{\dphidr}{\dxdy{\Phi}{r}}

@@

Following \citet{Quataert:2004a} (see also \citealt{HolzerAxford:1970a,De-ColleGuillochon+:2012a,ShcherbakovWong+:2014a}), we calculate the gas density profile $\rho$ of the CNM for each galaxy within our sample by solving the equations of one-dimensional (spherically symmetric) time-dependent hydrodynamics,
\begin{align}
  &\drhodt+\frac{1}{r^2}\frac{\partial}{\partial r}\left(\rho r^2 v\right)=q \label{eq:drhodt}\\
  &\rho \left(\frac{\partial v}{\partial t} + v\frac{\partial v}{\partial r}\right) =-\dpdr- \rho\frac{GM_{\bf enc}}{r^{2}} -q v \label{eq:dvdt}\\
  &\rho T\left(\frac{\partial s}{\partial t} + v\frac{\partial
      s}{\partial r}\right)=q\left[\ke+\kew-\gammaf \cs \right] +\nabla\cdot(\kappa \nabla T)
, 
\label{eq:dsdt}
\end{align}
where $v$, $T$, $p$ and $s$ are the radial velocity, temperature,
pressure and specific entropy of the gas, respectively, and $M_{\rm enc} = M_{\star}(r) + \Mbh$ is the
enclosed mass.  We assume an ideal gas equation of state with $p =
\rho kT/\mu m_p$ with $\mu = 0.62$  and $\gamma = 5/3$. {\bf BDM: take $\mu = 0.62$ here for fully ionized gas}

@@

The SMBH accretion rate is simply given by the integrated mass loss
rate interior to the stagnation radius (equation~[\ref{eq:q}]), 
%AG: Should we have separate cores and cusps? Or just write a general
%relationship in terms of gamma?
\begin{eqnarray}
  \dot{M} &=& \frac{\eta \Menc(\rs)}{\th} \nonumber \\
&\approx&
  \begin{cases}
    1.6 \times 10^{-3} M_{\bullet,8}^{1.9}
    \chi^{-4} v_{500}^{-4}  \eta \Msun \, \pyear& \text{core} \\
    3.4 \times 10^{-3} \Mbheight^{1.4} \chi^{-2}
    \chi^{-2} v_{500}^{-2}  \eta \Msun \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:mdot_analytic}
\end{eqnarray}

@@

For each solution we calculate the accretion rate $\dot{M}$ onto the
SMBH in units of the Eddington accretion rate $\dot{M}_{\rm edd}
\equiv L_{\rm edd}/0.1 c^{2}$, where $L_{\rm edd} \approx 1.4\times
10^{46}M_{\bullet,8}$ erg s$^{-1}$.  Figure~\ref{fig:mdot_mass} shows
$\dot{M}/\dot{M}_{\rm edd}$ as a function of $\Mbh$ for different
values of $\vwO =$ 150, 300 and 600 km s$^{-1}$.

Note that since we expect $\dot{M}\sim\vwO^{-2}$ for a cusp galaxy and
$\dot{M}\sim\vwO^{-4}$ for a core galaxy
(equation~[\ref{eq:mdot_analytic}]), the mass accretion rate is
expected to be a sensitive function of the assumed heating rate.
However, the accretion rate is proportional to the stellar mass
enclosed inside of $\rs$, so for core galaxies $\dot{M}\sim \Mbh
(\rs/\rsoi)^{2}$,while for cusp galaxies $\dot{M}\sim \Mbh
(\rs/\rsoi)$.  This steeper dependence is enough to offset the
tendency for core galaxies to have a larger $\rs/\rsoi$ for fixed
$\eta$ (see Figure~\ref{fig:stag}). Therefore, for $\rs/\rsoi<1$ core
galaxies would have a smaller $\dot{M}$ than cusp galaxies.  For
$\vwO=1000$ km s$^{-1}$, core galaxies have a smaller accretion rate
for a given SMBH mass.
%% AG-The variation in accretion rate for w/ vw for core galaxies is comparable to that of cusp galaxies and belies the steep dependence stated above.
%% AG-Note separation of cores and cusps (at vw=1000 km/s, where there are enough core galaxies for this comparison to be made). Also, note that the spread in core galaxies is far larger.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{mdot_mass.eps}
  \caption{\label{fig:mdot_mass}  Accretion rate $\dot{M}$ in units of the Eddington rate $\dot{M}_{\rm edd}$ as a function of SMBH mass for each galaxy in our sample, calculated for different values of the wind heating parameter $\vwO =$ 200 km s$^{-1}$ ({\it green}), 500 km s$^{-1}$ ({\it orange}), and 1000 km s$^{-1}$ ({\it blue}).  Squares correspond to cusp galaxies, while triangles correspond to cores.}
\end{figure}

@@

  \begin{enumerate}
  \item We find that $\eddr \propto \Mbh^{\alpha}$, where the value of
    $\alpha$ depends on the slope of the stellar density profile
    ($\alpha\simeq0.5,1$ for cusp and core galaxies respectively). One
    possible solution that the wind mass loss rate (i.e. $\eta$)
    decreases with $\Mbh$, as would be expected if low mass SMBH
    preferentially hosted older stellar populations. Another possible
    explanation is that the effective heating rate of the CNM $v_{\rm w,0}$
    decreases with decreasing SMBH mass. This could again be due to an
    older stellar population, since $v_{\rm w,0} \propto \eta^{−1/2}$
    for heating sources such as SN Ia and millisecond pulsars.
  \item We compare the $\Mdot$ for each of our solution to the one
    calculated from the Bondi formula. We find that they generally to
    a factor of a few. In our model the accretion rate is determined
    from the gas properties at the stagnation radius, $\rs$, which
    agrees with the Bondi radius, $\rb$, within a factor of a
    few. Thus, use of the Bondi formula is a good approximation.
  \item We compare our solutions for density and temperature to those
    inferred from Chandra x-ray observations by \citet{AllenDunn+:2006a}
    and \citet{RussellMcNamara+:2013a} for a handful of galaxies in
    which our sample overlap with theirs. In general, we cannot match
    the observed gas density slope on large scales. 
  \end{enumerate}

@@

In Fig. \ref{NickPlot} we plot $\dot{M}$, in arbitrary units, as a function of time, for three different stellar density profiles $\Gamma = \{1.1, 1.5, 1.9\}$ (which are normalized to have the same mass at an influence radius $r_{\rm soi}=10~{\rm pc}$ around a $10^7M_\odot$ SMBH).  We parametrize the wind velocity as 
\begin{equation}
\frac{v_{\rm w}}{\rm km~s^{-1}}=600-400\tanh\left( \frac{t-10^{7.5}~{\rm yr}}{10^{7}~{\rm yr}}\right).
\end{equation}
The ``impulsive burst'' mode of star formation produces large ($\sim 10$) differences between the three $\dot{M}$ curves at early times, when $r_{\rm s}$ is small, but smaller ($\sim 3$) differences at late times, when $r_{\rm s}$ is large.  We also plot, as dotted curves, a simple model for the ``continuous'' mode of star formation, where mass loss is calculated as $\bar{\eta} = \int\eta(t)dt/t_{\rm h} \approx 4$ and an average energy injection in the wind is calculated as $\bar{v_{\rm w}^2}=\int\eta(t)v_{\rm w}^2(t)dt/\bar{\eta} \approx (800~{\rm km~s}^{-1})^2$.  Interestingly, the continuous mode of star formation produces small differences from late-time $\dot{M}$ seen in cusp galaxies with impulsive star formation; however, continuous mode star formation decreases late-time $\dot{M}$ by an order of magnitude relative to impulsive star formation in core galaxies.
  \footnotesize{
    \bibliographystyle{mn2e}
    \bibliography{master}
  }
\end{document}

>>>>

\newcommand{\dxdy}[2]{\frac{d #1}{d #2} }
\newcommand{\ddr}[1]{\dxdy{#1}{r}}
\newcommand{\drhodt}{\dxdy{\rho}{t}}
\newcommand{\dpdr}{\dxdy{p}{r}}
\newcommand{\dvdr}{\dxdy{v}{r}}
\newcommand{\dsdr}{\dxdy{s}{r}}
\newcommand{\dphidr}{\dxdy{\Phi}{r}}

@@

Following \citet{Quataert:2004a} (see also \citealt{HolzerAxford:1970a,De-ColleGuillochon+:2012a,ShcherbakovWong+:2014a}), we calculate the gas density profile $\rho$ of the CNM for each galaxy within our sample by solving the equations of one-dimensional (spherically symmetric) time-dependent hydrodynamics,
\begin{align}
  &\drhodt+\frac{1}{r^2}\frac{\partial}{\partial r}\left(\rho r^2 v\right)=q \label{eq:drhodt}\\
  &\rho \left(\frac{\partial v}{\partial t} + v\frac{\partial
      v}{\partial r}\right) =-\frac{\partial p}{\partial r}- \rho\frac{GM_{\bf enc}}{r^{2}} -q v \label{eq:dvdt}\\
  &\rho T\left(\frac{\partial s}{\partial t} + v\frac{\partial
      s}{\partial r}\right)=q\left[\ke+\kew-\gammaf \cs \right] +\nabla\cdot(\kappa \nabla T)
, 
\label{eq:dsdt}
\end{align}
where $v$, $T$, $p$ and $s$ are the radial velocity, temperature,
pressure and specific entropy of the gas, respectively, and $M_{\rm enc} = M_{\star}(r) + \Mbh$ is the
enclosed mass.  We assume an ideal gas equation of state with $p =
\rho kT/\mu m_p$ with $\mu = 0.62$  and $\gamma = 5/3$. {\bf BDM: take $\mu = 0.62$ here for fully ionized gas}

@@

The SMBH accretion rate is simply given by the integrated mass loss
rate interior to the stagnation radius (equation~[\ref{eq:q}]), 
%AG: Should we have separate cores and cusps? Or just write a general

\begin{eqnarray}
  \dot{M} &=& \frac{\eta \Menc(\rs)}{\th} \nonumber \\
&\approx&
  \begin{cases}
    1.5 \times 10^{-3} M_{\bullet,8}^{1.76}
    \chi^{-3.8} v_{500}^{-3.8}  \eta \Msun \, \pyear& \text{core} \\
    2.6 \times 10^{-3} \Mbheight^{1.48} 
    \chi^{-2.4} v_{500}^{-2.4}  \eta \Msun \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:mdot_analytic}
\end{eqnarray}

The Eddington ratio $\eddr$ is given by

\begin{eqnarray}
\eddr&\approx&
  \begin{cases}
    6.6 \times 10^{-4} M_{\bullet,8}^{0.76}
    \chi^{-3.8} v_{500}^{-3.8}  \eta \Msun \, \pyear& \text{core} \\
    1.2 \times 10^{-3} \Mbheight^{0.48} 
    \chi^{-2.4} v_{500}^{-2.4}  \eta \Msun \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:eddr_analytic}
\end{eqnarray}

@@

For each solution we calculate the accretion rate $\dot{M}$ onto the
SMBH in units of the Eddington accretion rate $\dot{M}_{\rm edd}
\equiv L_{\rm edd}/0.1 c^{2}$, where $L_{\rm edd} \approx 1.4\times
10^{46}M_{\bullet,8}$ erg s$^{-1}$.  Figure~\ref{fig:mdot_mass} shows
$\dot{M}/\dot{M}_{\rm edd}$ as a function of $\Mbh$ for different
values of $\vwO =$ 150, 300 and 600 km s$^{-1}$, and for both core
($\gamma$=0.1) and cusp galaxies ($\gamma$=0.8).

Note that since we expect $\dot{M}\sim\vwO^{-2}$ for a cusp galaxy and
$\dot{M}\sim\vwO^{-4}$ for a core galaxy
(equation~[\ref{eq:mdot_analytic}]), the mass accretion rate is
expected to be a sensitive function of the assumed heating rate.
However, the accretion rate is proportional to the stellar mass
enclosed inside of $\rs$, so for core galaxies $\dot{M}\sim \Mbh
(\rs/\rsoi)^{2}$,while for cusp galaxies $\dot{M}\sim \Mbh
(\rs/\rsoi)$.  This steeper dependence is enough to offset the
tendency for core galaxies to have a larger $\rs/\rsoi$ for fixed
$\eta$ (see Figure~\ref{fig:stag}). Therefore, for $\rs/\rsoi<1$ core
galaxies would have a smaller $\dot{M}$ than cusp galaxies.  For
$\vwO=600$ km s$^{-1}$, core galaxies have a smaller accretion rate
for a given SMBH mass. On the other hand, if  $\rs/\rsoi>1$ (as would
be expected for small $\vwO$), core galaxies would have a larger
accretion rate. Thus, for $\vwO=150$ km/s, core galaxies have a large
accretion rate for a given $\Mbh$.  
%% AG-The variation in accretion rate for w/ vw for core galaxies is comparable to that of cusp galaxies and belies the steep dependence stated above.
%% AG-Note separation of cores and cusps (at vw=1000 km/s, where there are enough core galaxies for this comparison to be made). Also, note that the spread in core galaxies is far larger.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{mdot_mass.eps}
  \caption{\label{fig:mdot_mass} Accretion rate $\dot{M}$ in units of
    the Eddington rate $\dot{M}_{\rm edd}$ as a function of SMBH mass
    for each galaxy in our sample, calculated for different values of
    the wind heating parameter $\vwO =$ 150 km s$^{-1}$ ({\it green}),
    600 km s$^{-1}$ ({\it orange}), and 1000 km s$^{-1}$ ({\it blue}).
    Squares correspond to cusp galaxies ($\gamma=0.8$), while
    triangles correspond to cores ($\gamma$=0.1). The curves correspond
    to the analytic results for $\eddr$ from equation
    (\ref{eq:eddr_analytic}) for cusps (solid lines) and cores (dashed
    lines).  }
\end{figure}

@@

  \begin{enumerate}
  %Younger population=>higher accretion rate; generally higher eta and
  %lower vw except for very young stellar populations.
  \item We find that $\eddr \propto \Mbh^{\alpha}$, where the value of
    $\alpha$ depends on the slope of the stellar density profile
    ($\alpha\simeq0.5,1$ for cusp and core galaxies respectively). One
    possible solution that the wind mass loss rate (i.e. $\eta$)
    decreases with $\Mbh$, as would be expected if low mass SMBH
    preferentially hosted younger stellar populations. Another possible
    explanation is that the effective heating rate of the CNM $v_{\rm w,0}$
    decreases with decreasing SMBH mass. This could again be due to an
    older stellar population, since $v_{\rm w,0} \propto \eta^{−1/2}$
    for heating sources such as SN Ia and millisecond pulsars.
  \item We compare the $\Mdot$ for each of our solution to the one
    calculated from the Bondi formula. We find that they generally to
    a factor of a few. In our model the accretion rate is determined
    from the gas properties at the stagnation radius, $\rs$, which
    agrees with the Bondi radius, $\rb$, within a factor of a
    few. Thus, use of the Bondi formula is a good approximation.
  \item We compare our solutions for density and temperature to those
    inferred from Chandra x-ray observations by \citet{AllenDunn+:2006a}
    and \citet{RussellMcNamara+:2013a} for a handful of galaxies in
    which our sample overlap with theirs. In general, we cannot match
    the observed gas density slope on large scales. 
  \end{enumerate}

@@

In Fig. \ref{NickPlot} we plot $\dot{M}$, in arbitrary units, as a
function of time, for three different stellar density profiles $\Gamma
= \{1.1, 1.5, 1.9\}$ (which are normalized to have the same mass at an
influence radius $r
_{\rm soi}=10~{\rm pc}$ around a $10^7M_\odot$ SMBH).  We parametrize the wind velocity as 
\begin{equation}
\frac{v_{\rm w}}{\rm km~s^{-1}}=600-400\tanh\left( \frac{t-10^{7.5}~{\rm yr}}{10^{7}~{\rm yr}}\right).
\end{equation}
The ``impulsive burst'' mode of star formation produces large ($\sim 10$) differences between the three $\dot{M}$ curves at early times, when $r_{\rm s}$ is small, but smaller ($\sim 3$) differences at late times, when $r_{\rm s}$ is large.  We also plot, as dotted curves, a simple model for the ``continuous'' mode of star formation, where mass loss is calculated as $\bar{\eta} = \int\eta(t)dt/t_{\rm h} \approx 4$ and an average energy injection in the wind is calculated as $\bar{v_{\rm w}^2}=\int\eta(t)v_{\rm w}^2(t)dt/\bar{\eta} \approx (800~{\rm km~s}^{-1})^2$.  Interestingly, the continuous mode of star formation produces small differences from late-time $\dot{M}$ seen in cusp galaxies with impulsive star formation; however, continuous mode star formation decreases late-time $\dot{M}$ by an order of magnitude relative to impulsive star formation in core galaxies.
  \footnotesize{
    \bibliographystyle{mn2e}
    \bibliography{master}
  }
\end{document}
